---
title: "Data Cleaning Sample"
author: Janice Cho
output:
  html_document:
    toc: true
    toc_depth: 3
---

## Libraries
Load libraries.
```{r load_library, message = FALSE}
library(dplyr)
library(tidyr)
library(readxl)
library(gt)
library(gtExtras)
```


## Load Data

Load 9 waves of KLoSA data (light format) into 9 dataframes.
```{r load_data, include = FALSE}
all_waves <- list()
for (i in 1:9) {
  file_name <- paste0("lt_data/Lt0", i, ".xlsx")
  data_name <- paste0("wave_", i)
  all_waves[[i]] <- assign(data_name, read_excel(file_name))
}
```

## Clean Data

### Select variables of interest:

Display IDs and names of variables (for presentation purposes).
```{r var_table, echo = FALSE}
var_ids = c("pid",
            "hhid",
            "w0#A002_age",
            "w0#gender1",
            "w0#marital",
            "w0#hhsize",
            "w0#edu",
            "w0#wgt_c",
            "w0#wgt_p",
            "w0#pinc",
            "w0#present_ecotype",
            "w0#retired",
            "w0#G031",
            "w0#transferfrom",
            "w0#transferto",
            "w0#Ba003",
            "w0#hhnetassets")

var_labels = c("personal id",
               "household id",
               "age",
               "gender",
               "current marital status",
               "household size",
               "highest level of education",
               "cross-sectional weight",
               "longitudinal weight",
               "total personal income",
               "present economic activity status",
               "whether or not retired",
               "perceived socioeconomic status (starting from wave 2)",
               "total amount of financial support from\nchildren+parents+other family",
               "total amount of financial support offered\nto children+parents+other family",
               "number of alive children",
               "household net assets")

var_id_table <- data.frame(var_ids, var_labels) |> 
  gt() |> 
  gt_add_divider(
    var_ids,
    weight = px(0.75)
  ) |>
  tab_header(
    title = "Variables of Interest"
  ) |>
  tab_style(
    style = list(
      cell_fill(color = "#e3e3e3")
    ),
    locations = cells_title(groups = "title")
  ) |> 
  cols_label(
    var_ids = "ID",
    var_labels = "Variable"
  ) |> 
  cols_width(var_labels ~ px(200))

var_id_table
  
```

Gather names of variables of interest from each wave.
```{r get_var_names, include = FALSE}
# Function to assign variable names for each wave.
assign_var_names <- function(suffix, len = 9) {
  vars_list <- vector("character", length = len)
  for (i in 1:len) {
    vars_list[i] <- paste0("w0", i, suffix)
  }
  return(vars_list)
}

# Add variables missing from wave 1 to wave 1 and set everything to NA.
wave_1 <- wave_1 |> 
  mutate(w01G031 = NA,
         w01wgt_p = NA,
         w01hhnetassets = NA)
all_waves[[1]] <- wave_1

# Get variable names for each variable of interest.
cross_weight_vars <- assign_var_names(suffix = "wgt_c")
long_weight_vars <- assign_var_names(suffix = "wgt_p")
age_vars <- assign_var_names(suffix = "A002_age")
gender_vars <- assign_var_names(suffix = "gender1")
marital_vars <- assign_var_names(suffix = "marital")
hhsize_vars <- assign_var_names(suffix = "hhsize")
edu_vars <- assign_var_names(suffix = "edu")
pinc_vars <- assign_var_names(suffix = "pinc")
retired_vars <- assign_var_names(suffix = "retired")
ecotype_vars <- assign_var_names(suffix = "present_ecotype")
perceived_SES_vars <- assign_var_names(suffix = "G031")
transfer_from_vars <- assign_var_names(suffix = "transferfrom")
transfer_to_vars <- assign_var_names(suffix = "transferto")
alive_child_vars <- assign_var_names(suffix = "Ba003")
hh_asset_vars <- assign_var_names(suffix = "hhnetassets")

# Rename wave 9 "HHID" to "hhid".
if (!("hhid" %in% colnames(wave_9))) {
  wave_9 <- wave_9 |> 
    rename(hhid = HHID)
  all_waves[[9]] <- wave_9
}
```

Select variables of interest in each wave.
```{r select_waves, echo = FALSE}
# Create select_waves list to hold dataframes of clean waves.
select_waves <- list()

# Create vector of years corresponding to waves.
years <- seq.int(from = 2006, to = 2022, by = 2)

# Function to select the same variables of interest from each dataframe.
select_vars <- function(i, vars_of_interest) {
  vars_of_interest <- c("pid", "hhid", vars_of_interest)
  
  select_df <- all_waves[[i]] |> 
    select(all_of(vars_of_interest)) |> 
    mutate(wave = i,
           year = years[i]) |> 
    select(wave, year, everything()) |> 
    rename_with(~ gsub("w0\\d{1}", "", .x), starts_with("w0")) |> 
    rename(
      age = A002_age,
      gender = gender1,
      perceived_SES = G031,
      alive_child = Ba003,
      ecotype = present_ecotype
    )
  
  return(select_df)
}

# Select variables of interest in each wave and store in select_waves.
for (i in 1:9) {
  vars_of_interest <- c(cross_weight_vars[i],
                        long_weight_vars[i],
                        age_vars[i],
                        gender_vars[i],
                        marital_vars[i],
                        hhsize_vars[i],
                        edu_vars[i],
                        pinc_vars[i],
                        retired_vars[i],
                        ecotype_vars[i],
                        perceived_SES_vars[i],
                        transfer_from_vars[i],
                        transfer_to_vars[i],
                        alive_child_vars[i],
                        hh_asset_vars[i])
  
  select_waves[[i]] <- select_vars(i, vars_of_interest)
}

head(select_waves[[9]])
```

Create single dataframe to hold all waves.
```{r combine_select_waves, include = FALSE}
all_select_waves <- bind_rows(select_waves)
head(all_select_waves, 10)
```

### Recode variables:

Recode variables where necessary.
```{r recode_variables}
# Create vectors of factor levels.
perceived_SES_levels = c("Low", "Middle", "High")
marital_levels = c("Married", "Widowed", "Divorced/Separated", "Single")
edu_levels = c("Less than high school", "High school", "College or above")
ecotype_levels = c("Employed", "Unemployed", "Economically inactive")
age_levels = c("45-54", "55-64", "65-74", "75-84", "85+")

# Recode variables.
all_clean <- all_select_waves |> 
  mutate(
    # Turn "I don't know" and "Refuse to Answer" into NA for personal income.
    pinc = if_else(pinc < 0, NA, pinc),
    
    # Turn perceived_SES into factor.
    perceived_SES = case_when(
      perceived_SES %in% 5:6 ~ "Low",
      perceived_SES %in% 3:4 ~ "Middle",
      perceived_SES %in% 1:2 ~ "High",
      .default = NA
    ),
    perceived_SES = factor(perceived_SES, levels = perceived_SES_levels),
    
    # Turn marital into factor.
    marital = case_when(
      marital == 1 ~ "Married",
      marital == 4 ~ "Widowed",
      marital %in% 2:3 ~ "Divorced/Separated",
      marital == 5 ~ "Single"
    ),
    marital = factor(marital, levels = marital_levels),
    
    # Turn education level into factor.
    edu = case_when(
      edu <= 2 ~ "Less than high school",
      edu == 3 ~ "High school",
      edu >= 4 ~ "College or above"),
    edu = factor(edu, levels = edu_levels),
    
    # Turn economic activity status into factor.
    ecotype = case_when(
      ecotype == 1 ~ "Employed",
      ecotype == 2 ~ "Unemployed",
      ecotype == 3 ~ "Economically inactive"),
    ecotype = factor(ecotype, levels = ecotype_levels),
    
    # Turn gender into factor.
    gender = if_else(gender == 1, "Male", "Female"),
    gender = factor(gender, levels = c("Male", "Female")),
    
    # Add age group category.
    age_cat = case_when(
      age < 55 ~ "45-54",
      age %in% 55:64 ~ "55-64",
      age %in% 65:74 ~ "65-74",
      age %in% 75:84 ~ "75-84",
      age >= 85 ~ "85+"
    ),
    age_cat = factor(age_cat, levels = age_levels)
  )

head(all_clean |> filter(wave == 2), 10)
```